---
title: "Aides Drug"
author: "Shawn Olson"
date: "09/26/24"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
```{r}
Female<- filter(triUmp, gender=="F")
```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected + treatment, data=Female)
barchartGC(~infected + treatment, data=Female, type="percent")

```

The visual graph shows us that in a group that are only female; the ones that took the drug have a higher chance getting aids than the ones that took the placebo who didn't take the drug at all.



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
table1 <- xtabs (~infected + treatment,data=Female)
rowPerc(table1)
colPerc(table1)
```

in the visual numerical graph results the percentage in the group that didnt get aids We can see that the percentage in the group that didnt get aids are pretty close to 41% when you look at the people who got aids are pretty close to 60%

```{r}
chisq.test (table1)
chisqtestGC(table1)
fisher.test(table1)
```

There were three statistical test ran and each test gave us a probability with the results from before
with the group of female and their chances of getting aids with a drug or placebo. The first test had a
probability of 0.00007686, the second probability is 0.0001 and the third's probability is 0.00006596.


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

The probabilities are the statistical differences made between the female centered groups and the
drug or placebo they took that would tell them if the drug prevents aids.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

I graphed the results of a group that is only of female and the results showed us whether the drug they
were given helped prevent aids. The bar graph then showed us that the female given the drug has a
higher chance of getting aids that the female that were given a placebo. I then ran a numerical graph
and the result showed me that the female that got aids were a higher 20% than the female that didn't
get aids. I then also ran three statistical test that told me the probability of a female getting aids from
the drug that was given in the test. From the results of all the test ran the drug actually help the female
in the group get aids then if they were given a placebo without. We can conclude from this test that the
drug did not work with females.
 


